### 2016-02-13 paper/1-introduction.tex
Apply reviewers comments.

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
And while \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, this is often insufficient.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to \textsf{DEVS} simulation.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other in specific models.
The applicability of parallel simulation capabilities of current tools is therefore limited.

Users who simulate a wide variety of models, with different ideal synchronization protocols, are therefore out of luck:
either they accept lower performance for some models, or they use different simulation kernels.
Neither is acceptable for complex models: performance can decrease to unacceptable levels, and simulation kernel APIs can diverge significantly.

This paper introduces DEVS-Ex-Machina (``dxex''), our simulation tool which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparant to the simulated model.
Users should merely determine, at the start of simulation, which protocol they wish to use.
Our tool is based on PythonPDEVS, but implemented in C++11, to increase both performance and portability across different platforms.

>>>>

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
And while \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, this is often insufficient.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to \textsf{DEVS} simulation.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other in specific models.
The applicability of parallel simulation capabilities of current tools is therefore limited.
% Paragraph moved to beneath intro dxex, else we can't refer to our tool
This paper introduces DEVS-Ex-Machina (``dxex''), our simulation tool which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparent to the simulated model.
Users should merely determine, at the start of simulation, which protocol they wish to use.
Our tool is based on PythonPDEVS, but implemented in C++11, to increase both performance and portability across different platforms. Unlike PythonPDEVS dxex only supports multicore parallelism.

Users who simulate a wide variety of models, with different ideal synchronization protocols, are therefore in luck:
with our tool the model implementation can be reused with a different protocol to test which offers better performance.

### 2016-02-13 paper/3-features.tex
Apply reviewers comments.

We will not make a detailed comparison with PythonPDEVS here, but only mention some supported features.
Dxex supports, similarly to PythonPDEVS, the following features: direct connection~\cite{SymbolicFlattening}, \textsf{Dynamic Structure DEVS}~\cite{DSDEVS}, termination conditions, and a modular tracing and scheduling framework~\cite{PythonPDEVS}.
But whereas PythonPDEVS only supports optimistic synchronization, dxex support multiple synchronization protocols (though only in parallel).
This is in line with the design principle of PythonPDEVS: allow users to pass performance hints to the simulation kernel.
In our case, a user can pass the simulation kernel the ``hint'' as to which synchronization protocol must be used for this model.
Our implementation in C++11 also allows for optimizations which were plainly impossible in an interpreted language.

@@

Always defining a lookahead function might seem redundant, especially if users will never use conservative synchronization.
Especially since defining the lookahead is often non-trivial and dependent on intricate model details.
The more attractive option is for the simulation tool to provide a default lookahead function, such that simulation can run anyway, but maybe not at peak performance.
Depending on the model, simulation performance might still be faster than sequential simulation.

Defining a lookahead function is therefore recommended in combination with conservative synchronization, but is not a necessity, as a default can be used otherwise.

>>>>

We will not make a detailed comparison with PythonPDEVS here, but only mention some supported features.
Dxex supports, similarly to PythonPDEVS, the following features: direct connection~\cite{SymbolicFlattening}, \textsf{Dynamic Structure DEVS}~\cite{DSDEVS}, termination conditions, and a modular tracing and scheduling framework~\cite{PythonPDEVS}.
But whereas PythonPDEVS only supports optimistic synchronization, dxex support multiple synchronization protocols (though only in parallel).
This is in line with the design principle of PythonPDEVS: allow users to pass performance hints to the simulation kernel.
In our case, a user can pass the simulation kernel sets the synchronization protocol to be used for this model.
Our implementation in C++11 also allows for optimizations which were plainly impossible in an interpreted language.

@@

Always defining a lookahead function might seem redundant, especially if users will never use conservative synchronization.
Especially since defining the lookahead is often non-trivial and dependent on intricate model details.
The more attractive option is for the simulation tool to provide a default lookahead function, such that simulation can run anyway, but maybe not at peak performance.
%Depending on the model, simulation performance might still be faster than sequential simulation.

Defining a lookahead function is therefore recommended in combination with conservative synchronization, but is not a necessity, as a default of $epsilon$ can be used otherwise.

### 2016-02-13 paper/4-performance.tex
EPS files added + R script changes + Dataset to obtain figures added
Apply reviewers comments.

For all benchmarks, results are well within a 5\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3 -flto}'').
To guarantee comparable results, no I/O was performed during benchmarks.
Before benchmarking, simulation traces were used to verify that adevs and dxex return exactly the same simulation results.
Benchmarks were performed using Linux, but our simulation tool works equally well on Windows and Mac.

@@

\subsubsection{Queue}
In the Queue model, we increase both width and depth simultaneously.
For example, the $400$ models configuration is obtained with a width and depth of $20$.
As can be seen in Figure~\ref{fig:Queue_benchmark}, dxex considerably outperforms adevs.
Through profiling, we determined that adevs spends much time on the handling of simulation messages, while this is avoided in dxex's simulation algorithm.
Both simulation tools have similar complexities, though dxex is much faster thanks to its more efficient simulation control algorithms.

@@

\subsubsection{HighInterconnect}
In the HighInterconnect model, we increase the number of atomic models, thus quadratically increasing the number of couplings and the number of external transitions.
As can be seen in Figure~\ref{fig:Interconnect_benchmark}, adevs outperforms dxex by a fair margin.
Analysis showed that this is caused by the high amount of exchanged events: event creation is much slower in dxex than it is in adevs, despite dxex's use of memory pools.

@@

Similar to the sequential benchmarks, Figure~\ref{fig:queue_benchmark_parallel} shows that dxex outperforms adevs, but now in terms of speedup.
Results indicate that our implementation of conservative synchronization achieves much higher speedups than adevs.
This simulation is the ideal case for our conservative implementation, approaching near linear speedup.
%TODO: Why is this the ideal case?
Conservative synchronization also seems to be better than optimistic synchronization in this case, at the cost of providing the lookahead.

@@

\subsubsection{PHold}
In the Phold model, we first investigate the influence of the fraction of remote events on the speedup.
When remote events are rare, optimistic synchronization rarely has to roll back, thus increasing performance.
With more common remote events, however, optimistic synchronization quickly slows down due to frequent rollbacks.
Conservative synchronization, on the other hand, is mostly unconcerned with the number of remote events: the mere fact that a remote event can happen, causes it to block.
Even though a single synchronization protocol is always ideal in this case, it already shows that different synchronization protocols respond differently to a changing model.
Adevs is again significantly slower during conservative synchronization.
Analysis of profiling results shows that exception handling in adevs is the main cause.

\begin{figure}
    \center
    \includegraphics[width=\plotfraction\columnwidth]{fig/phold_remotes.eps}
    \caption{Phold benchmark results for parallel simulation using six cores, with varying fraction of remote events.}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\plotfraction\columnwidth]{fig/phold_priority.eps}
    \caption{Phold benchmark results for parallel simulation using six cores, with varying amount of high-priority events.}
    \label{fig:phold_priority}
\end{figure}

>>>>

For all benchmarks, results are well within a 5\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3 -flto}'').
To guarantee comparable results, no I/O was performed during benchmarks.
Before benchmarking, simulation traces were used to verify that adevs and dxex return exactly the same simulation results.
Benchmarks were performed using Linux, but our simulation tool works equally well on Windows and Mac. The exact parameters for each benchmark can be found in the repository, as well as the data used in this paper.

@@

\subsubsection{Queue}
In the Queue model, we increase both width and depth simultaneously.
For example, the $400$ models configuration is obtained with a width and depth of $20$.
As can be seen in Figure~\ref{fig:Queue_benchmark}, dxex considerably outperforms adevs.
Through profiling, we determined that adevs spends much time on the routing of simulation messages through the hierarchy, while dxex avoids this by flattening the model structure before the simulation starts.
Both simulation tools have similar complexities, though dxex is much faster thanks to its more efficient simulation control algorithms.

@@

\subsubsection{HighInterconnect}
In the HighInterconnect model, we increase the number of atomic models, thus quadratically increasing the number of couplings and the number of external transitions.
As can be seen in Figure~\ref{fig:Interconnect_benchmark}, adevs outperforms dxex by a fair margin.
Analysis showed that this is caused by the high amount of exchanged events: event creation is much slower in dxex than it is in adevs, despite dxex's use of memory pools. For the speedup plot the amount of models is kept constant at 8.

@@

Similar to the sequential benchmarks, Figure~\ref{fig:queue_benchmark_parallel} shows that dxex outperforms adevs, but now in terms of speedup.
Results indicate that our implementation of conservative synchronization achieves much higher speedups than adevs.
This simulation is the ideal case for our conservative implementation, approaching near linear speedup, since there are no dependency cycles between models.
Conservative synchronization also seems to be better than optimistic synchronization in this case, at the cost of providing the lookahead.

@@

\subsubsection{PHold}
In the Phold model, we first investigate the influence of the fraction of remote events on the speedup.
When remote events are rare, optimistic synchronization rarely has to roll back, thus increasing performance.
With more common remote events, however, optimistic synchronization quickly slows down due to frequent rollbacks.
Conservative synchronization, on the other hand, is mostly unconcerned with the number of remote events: the mere fact that a remote event can happen, causes it to block.
Even though a single synchronization protocol is always ideal in this case, it already shows that different synchronization protocols respond differently to a changing model.
Adevs is again significantly slower during conservative synchronization.
Analysis of profiling results shows that exception handling in adevs is the main cause. 
% Addin ? To keep the models equivalent, the adevs version does not provide the \{begin,end\}Lookahead methods, which accounts for the exception handling.

\begin{figure}
    \center
    \includegraphics[width=\plotfraction\columnwidth]{fig/phold_speedup_remotes.eps}
    \caption{Phold benchmark results for parallel simulation using six cores, 4 atomics per node, with varying fraction of remote events.}
\end{figure}

@@

\begin{figure}
    \center
    \includegraphics[width=\plotfraction\columnwidth]{fig/phold_speedup_priority.eps}
    \caption{Phold benchmark results for parallel simulation using six cores, with varying amount of high-priority events.}
    \label{fig:phold_priority}
\end{figure}

### 2016-02-13 paper/5-related-work.tex
Apply reviewers comments.

Previous \textsf{DEVS} simulation tools have already implemented multiple synchronization protocols, though mostly in an ad-hoc way.
For example CD++~\cite{CD++} has both a conservative (CCD++~\cite{CCD++}) and optimistic (PCD++)~\cite{PCD++}) variant.
Despite the implementation of both protocols, they are different projects entirely, and are incompatible with modern compilers.
Dxex, on the other hand, is a single project, where switching between different synchronization protocols is as simple as switching any other configuration parameter.
CD++, however, implements both conservative and optimistic synchronization for distributed simulation, whereas we limit ourselves to parallel simulation.
By limiting our approach to parallel simulation, we are able to achieve higher speedups through the use of shared memory communication.

>>>>

Previous \textsf{DEVS} simulation tools have already implemented multiple synchronization protocols, though none have done it in a strictly modular way that allows straightforward protocol switching for a single given model.
For example CD++~\cite{CD++} has both a conservative (CCD++~\cite{CCD++}) and optimistic (PCD++)~\cite{PCD++}) variant.
Despite the implementation of both protocols, they are different projects entirely, and are incompatible with modern compilers.
Dxex, on the other hand, is a single project, where switching between different synchronization protocols is as simple as switching any other configuration parameter.
CD++, however, implements both conservative and optimistic synchronization for distributed simulation, whereas we limit ourselves to parallel simulation.
By limiting our approach to parallel simulation, we are able to achieve higher speedups through the use of shared memory communication.
